Automated discovery \\ of interpretable gravitational-wave population models

### 2022-07-11 src/tex/ms.tex
Space out rows of table
Start discussing nested constraints

\begin{table*}[hbt!]
    \begin{center}
    \begin{tabular}{ l l l l }
    \hline
    \hline
    Complexity &  Score & Loss & Expression\\
    \hline
    \hline
    16 & 0.126   & 0.212  & $3.72 \rm{Cond}{\left(M - 9.27,0.9^{M} \right)} + 3.72 \rm{Gauss}{\left(0.52 M - 4.85 \right)}$\\
    25 & 0.501   & 0.0352 & $3.89 \left(\rm{Gauss}{\left(0.19 M - 6.54 \right)} + 0.45\right) \rm{Cond}{\left(M - 9.12,0.91^{M} \right)}$\\
    & & & $ + 3.89 \rm{Gauss}{\left(0.5 M - 4.8 \right)}$\\
    48 & 0.00771 & 0.0108 & $\left(\rm{Cond}{\left(M - 9.42,0.62 \cdot 0.9^{M} \right)} + 1.44 \rm{Gauss}{\left(0.51 M - 4.88 \right)}\right)$\\
    & & & $\left(5.11 \rm{Gauss}{\left(0.06 M - 4.67 \right)} + 7.82 \rm{Gauss}{\left(0.17 M - 5.86 \right)} + 3.26\right)$\\
    \hline
    \hline
    \end{tabular}
    \caption{Expressions obtained through symbolic regression with \pysr.
    In the search we perform, there are 30 equations with different complexities.
    We select three most representative equations by setting loss thresholds, and selecting the highest scoring expression below each threshold. This strategy rules out trivial expressions with low accuracy which had high scores only because they were adjacent to other trivial expressions.}
    \label{tab:equations}
    \end{center}
\end{table*}

@@

Here, we use symbolic regression for the first time to learn population density models in astrophysics.
One of the main advantages of symbolic regression is to learn a model into an interpretable language, and, therefore, we define a set of operators which frequently occur in GW population models: addition, subtraction, division, power laws, Gaussians (unary operator: $f(x)=\exp(-x^2)$), and conditional statements (binary operator: $g(x, y)=\operatorname{if}(x \geq 0, y, 0)$). 
% \todo{Specify how combinations of operators are ruled out}

@@

>>>>

\begin{table*}[hbt!]
    \begin{center}
    \begin{tabular}{ l l l l }
    \hline
    \hline
    Complexity &  Score & Loss & Expression\\
    \hline
    \hline
    16 & 0.126   & 0.212  & $3.72 \rm{Cond}{\left(M - 9.27,0.9^{M} \right)} + 3.72 \rm{Gauss}{\left(0.52 M - 4.85 \right)}$\\[0.4em]
    % Add more space between rows of table^
    25 & 0.501   & 0.0352 & $3.89 \left(\rm{Gauss}{\left(0.19 M - 6.54 \right)} + 0.45\right) \rm{Cond}{\left(M - 9.12,0.91^{M} \right)}$\\
    & & & $ + 3.89 \rm{Gauss}{\left(0.5 M - 4.8 \right)}$\\[0.4em]
    48 & 0.00771 & 0.0108 & $\left(\rm{Cond}{\left(M - 9.42,0.62 \cdot 0.9^{M} \right)} + 1.44 \rm{Gauss}{\left(0.51 M - 4.88 \right)}\right)$\\
    & & & $\left(5.11 \rm{Gauss}{\left(0.06 M - 4.67 \right)} + 7.82 \rm{Gauss}{\left(0.17 M - 5.86 \right)} + 3.26\right)$\\
    \hline
    \hline
    \end{tabular}
    \caption{Expressions obtained through symbolic regression with \pysr.
    In the search we perform, there are 30 equations with different complexities.
    We select three most representative equations by setting loss thresholds, and selecting the highest scoring expression below each threshold. This strategy rules out trivial expressions with low accuracy which had high scores only because they were adjacent to other trivial expressions.}
    \label{tab:equations}
    \end{center}
\end{table*}

@@

Here, we use symbolic regression for the first time to learn population density models in astrophysics.
One of the main advantages of symbolic regression is to learn a model into an interpretable language, and, therefore, we define a set of operators which frequently occur in GW population models: addition, subtraction, division, power laws, square, Gaussians (unary operator: $f(x)=\exp(-x^2)$), and conditional statements (binary operator: $g(x, y)=\operatorname{if}(x \geq 0, y, 0)$). 
To simplify our assumptions, all operators are assigned the same complexity of $1$.
We also introduce additional constraints on how operators may be nested inside each other to rule out uninterpretable expressions; these are defined in \cref{sec:constraints}.

% \todo{Specify how combinations of operators are ruled out}

@@

\appendix
\section{Operator constraints}
\label{sec:constraints}

We implement several constraints on the space of expressions, to rule out uninterpretable combinations of operators.
We impose the following nested constraints:
\begin{enumerate}
    \item No operator with arity of 1 (i.e., taking a single argument) may be nested with another arity-1 operator, unless the operator is square.
    \begin{enumerate}
        \item However, a square can only be nested inside itself once.
    \end{enumerate}
    \item A division operator may only be nested inside itself once.
    \item A power operator may not contain any arity-1 operator, except a square.
\end{enumerate}
